---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
library(tidyverse)
library(ggplot2)
activity <- read_csv("activity.zip")
```

## What is mean total number of steps taken per day?
```{r}
by_day <- group_by(activity, date)
day_tot <- summarize(by_day, sum_steps = sum(steps, na.rm = TRUE))
summary(day_tot$sum_steps)
```

The mean number of steps per day is `r round(mean(day_tot$sum_steps), 1)`, and
the median is `r median(day_tot$sum_steps)`.

Here's the histogram of the total number of steps per day.

```{r}
ggplot(day_tot, aes(sum_steps)) +
    geom_histogram(bins = 20, colour = "darkblue", fill = "lightblue") +
    labs(x = "Number of steps",
         title = "Total Number of Steps Taken Each Day")
```

## What is the average daily activity pattern?
```{r}
by_interval <- group_by(activity, interval)
intrvl_avg <- summarize(by_interval, avg_steps = mean(steps, na.rm = TRUE))
```

```{r}
ggplot(intrvl_avg, aes(x = interval, y = avg_steps)) +   
    geom_line() +  
    labs(title = "Average Daily Activity Pattern",
         x = "Interval",
         y = "Average number of steps per interval")
```

```{r}
intrvl_max <- intrvl_avg[which.max(intrvl_avg$avg_steps), ]
```

The interval that, on average, contains the maximum number of steps is the 
`r intrvl_max[1]`<sup>th</sup>, with `r round(intrvl_max[2], 0)` steps.

## Imputing missing values
```{r}
missing <- sum(is.na(activity$steps))
```

The total number of missing values in the dataset is `r missing`.

Missing values are replaced by the average for that interval.

```{r}
activity_imput <- transform(activity,
                            steps = ifelse(is.na(activity$steps),
                                           intrvl_avg$avg_steps[match(activity$interval,
                                                                      intrvl_avg$interval)],
                                           activity$steps))
```

```{r}
by_day_imput <- group_by(activity_imput, date)
day_tot_imput <- summarize(by_day, sum_steps = sum(steps, na.rm = TRUE))
ggplot(day_tot_imput, aes(sum_steps)) +
    geom_histogram(bins = 20, colour = "darkblue", fill = "lightblue") +
    labs(x = "Number of steps",
         title = "Total Number of Steps Taken Each Day",
         subtitle = "After imputation of missing values")
```


## Are there differences in activity patterns between weekdays and weekends?
